\section{Study on the Influence of Key Physical Parameters}\label{sec:parameter-study}

The simulation of an asymmetrical two-particle-contact involves a variety of physical geometry and material parameters which highly influence the sintering behavior and macroscopic response of the system to be observed.
A widely used approach to characterize a system is by it's dimensionless parameters.
The following study investigates the influence of the dimensionless parameters present in an asymmetrical two-particle-system.

\subsection{Construction of the Study}

The shape of a particle is here modeled as an ellipse an parameterized with the equivalent circular radius $\Radius_{\Particle}$ and the ovality $\Ovality \in [ 1, \infty ) $ (as the standard excentricity is an hardly illustrative value for the current purpose).
The local polar radius coordinate of a node is then given by \autoref{eq:parameter_study_ellipse_function}, with the main radii defined by $a = \Radius_{\Particle} \sqrt{\Ovality}$ and $b = \Radius_{\Particle} / \sqrt{\Ovality}$ with $a/b = o$.

\begin{equation}
    \Radius(\Angle) = \frac{a b}{\sqrt{\left( a \sin \Angle \right)^2 + \left( b \cos \Angle \right)^2  }}
    \label{eq:parameter_study_ellipse_function}
\end{equation}

\begin{table}
    \caption{Investigated Dimensionless Parameters and their Range}
    \label{tab:parameter_study_conditions}
    \begin{tblr}{colspec={lrrlX}, row{1}={l}}
        \toprule
        Parameter & Min & Max & Scale & remarks \\
        \midrule
        $\Radius_1 / \Radius_2$ & \num{1} & \num{10} & lin & equal material\\
        $\DiffusionCoefficient_{\GrainBoundary} / \DiffusionCoefficient_{\Surface}$  & \num{0.01} & \num{1} & log & equal material\\
        $\InterfaceEnergy_{\GrainBoundary} / \InterfaceEnergy_{\Surface}$ & \num{0.1} & \num{0.9} & lin & equal material \\
        $\DiffusionCoefficient_{\Surface2} / \DiffusionCoefficient_{\Surface1} $ & \num{1} & \num{100} & log & $\InterfaceEnergy_{\GrainBoundary} / \InterfaceEnergy_{\Surface1} = \num{0.5}$, $\DiffusionCoefficient_{\GrainBoundary} / \DiffusionCoefficient_{\Surface1} = \num{1}$ \\
        $\InterfaceEnergy_{\Surface2} / \InterfaceEnergy_{\Surface1}$ & \num{1} & \num{10} & lin & $\InterfaceEnergy_{\GrainBoundary} / \InterfaceEnergy_{\Surface1} = \num{0.5}$, $\DiffusionCoefficient_{\GrainBoundary} / \DiffusionCoefficient_{\Surface1} = \num{1}$ \\
        $o$ & \num{1} & \num{6} & lin & tip/tip contact, equal material \\
        $o$ & \num{1} & \num{6} & lin & tip/flank contact, equal material \\
        $o$ & \num{1} & \num{6} & lin & flank/flank contact, equal material \\
        \bottomrule
    \end{tblr}
\end{table}

\autoref{tab:parameter_study_conditions} lists the varied dimensionless parameters and their respective value ranges, where the indices 1 and 2 denote the first resp. the second particle.
The study includes:
\begin{description}
    \item[symmetrical contacts] with varied ratios of interface energies and diffusion coefficients between surface and grain boundary
    \item[asymmetrical geometry contacts] with circular particles of different sizes and oval particles in different contact modes
    \item[asymmetrical material contacts] with varied ratios of interface energies and diffusion coefficients on both sides
\end{description}

The initial contact of the particles is always taken in that way, that the distance of the particles equals \qty{1}{\percent} of the reference particle's radius $\Radius_1$.
Then, the intersection of the particle's is calculated and the grain boundary is placed as a straight line between the surface line intersection points.

\subsection{Discussion of the Results}

The results will be displayed as isoline-plots for shrinkage and neck size with the normalized time on the horizontal axis and the regarded dimensionless parameter on the vertical axis.

\input{tex/physical_behavior/parameter_study/particle_size_ratio}
\input{tex/physical_behavior/parameter_study/surface_boundary_diffusion_ratio}
\input{tex/physical_behavior/parameter_study/surface_boundary_energy_ratio}
\input{tex/physical_behavior/parameter_study/asymmetric_surface_diffusion_ratio}
\input{tex/physical_behavior/parameter_study/asymmetric_surface_energy_ratio}
