\chapter{Numerical Methods for Sintering Simulation}\label{ch:numerical-methods}

Sintering models can be distinguished in two main branches: macroscale and microscale models. 
The former regard the porous worpiece material as a continuum with variable density. 
Stresses and strains are usually modelled using visco-plastic material laws or similar while the density evolves during thermal treatment or plastic deformation (mechanical compaction). 
Numerical solution are commonly obtained by use of the \gls{FEM}.
The focus of these models usually lies in the analysis of density, strain and stress distributions, as well as damage estimation, in a component with often complex geometry.
Examples for these models are found in \textcite{Olevsky1998, Safonov2019, Kraft2004, Al-Qudsi2014, Shi2023}. 
These models will not be considered in the following, as this work concentrates on the second kind.

Microscale models regard the interaction between distinct grains resp. powder particles and the evolution of their interfaces to each other and the environment. 
The solution space typically includes from two to a limited number of distinct particles or a defined \gls{RVE}. 
Physics-based models model the distinct diffusion mechanisms occuring and their effect on the particle and/or pore shape. 

There are basically two ways of representing the interfaces of particles to each other or the environment: sharp and diffuse interfaces. 
Both were reported to be used in sintering simulation before.

A sharp interface means that there is a defined line were the interface is supposed to be located, on one side the one phase, on ther other side the other, without any transition.
Sharp interfaces are the most straightforward approach, as most people are used to think in that way, although in reality, interfaces have a non-zero width, typically in the scale of a few atom diameters. 
One may think about the adsorption and diffusion layer on solid surfaces or the mislalignment voids between crystallites in grain boundaries. 
However, the width of interfaces is usually much smaller then the size of particles or microscopic surface features, and can therefore be neglected.
Sharp interfaces can be described directly by the coordinates of their line string, either as a linear or non-linear spline or similar, or, by following an equality condition in the solution space (e.g. level-set method or cellular automata).
The curvature of such an interface can directly be calculated by interpolation of its coordinates, f.e. for use in \autoref{eq:stress-surface-curvature}.
Models of this type are discussed in detail in \autoref{sec:sharp-interfaces}. 

A diffuse interface means that there is a certain region of transition between the two phases.
The most common approach for that is the phase field method, where the phase present at a point is defined by the value of one or multiple auxiliary variables, whose values go smoothly between \num{0} and \num{1}.
For illustration purposes the interface is often defined at a ceratin threshold value of the phase field variables, but the model usually does not work with such, but only with the gradient of the variable.
There are no hard transitions in such models, but everything is smooth (except for the limitations of numerical discretization).
Models of this type are discussed in detail in \autoref{sec:diffuse-interfaces}.

\input{tex/numerical_methods/interfaces}
\input{tex/numerical_methods/monte_carlo}
\input{tex/numerical_methods/extremal_principle}
\input{tex/numerical_methods/inverse_problems}
