\section{Choice of Variables}\label{sec:variables}

First, the internal state variables $\Vect\InternalStateVariable$, the internal state velocities $\Vect\InternalStateVelocity$ and the fluxes $\Vect\Flux$ must be chosen.
The feasibility of the approach heavily depends on this choice.

The choice of the fluxes is straightforward with the diffusional fluxes.
One has two diffusional fluxes from each node $\Flux_{\Upper}$ and $\Flux_{\Lower}$ to the upper respectively the lower neighbor, however the flux of the upper node to the lower and the flux from the lower to the upper are always equal due to constancy of mass.
So in the following only $\Flux_{\Upper}$ shall be regarded and $\Flux_{\Lower}$ is substituted by $-\Flux_{\Upper}$ where necessary.

The internal state variables are the polar coordinates of the nodes $\left[ \Radius, \Angle \right]$ with pole in their particle's center, since they determine the thermodynamic forces and the main aim of the simulation is to follow the time evolution of the particles and nodes.
The coordinates of the particles $\left[ \X_{\Particle}, \Y_{\Particle} \right]$ are not used as internal state variables, since they are unambiguously defined, if the coordinates of all nodes relative to their particle are known and also which nodes are in contact to each other.
Moreover, the velocities of particle coordinates $\dot\X_{\Particle}$ and $\dot\Y_{\Particle}$ are auxiliary variables (included in $\Vect\AuxiliaryVariable$) to ease formulation of the geometric constraints.

For numerical and equation formulation reasons, the internal state velocities $\dot\Radius$ and $\dot\Angle$ are replaced by the velocities of node shift along normal and tangential surface vectors $\dot\Shift_{\Normal}$ and $\dot\Shift_{\Tangential}$.
As they are determined, translation into new coordinates is directly possible using the relations obtained in the following sections.

The external state variables (such as temperature $\Temperature$ and pressure $\Pressure$) are not considered in the following, since they are assumed constant.
